Molecular Gas Kinematics and Spectral Energy Distribution Properties of the Strongly-Lensed Quasar Host Galaxy RXS J1131$-$1231

### 2016-12-08 WRITEUP/ms.tex
after proof-reading by myself, will give to DR

We detect 2.2\,mm continuum emission at an
 integrated flux density of
1.2\pmm0.2 mJy, with a peak flux of
$S_{\nu}$\,=\,799\pmm88\,$\mu$Jy\,beam\pmOne
centered on the lensing galaxy (\Fig{cont}).
% We note that the reported flux uncertainties do not include those from absolute flux calibration.
Slightly extended emission
along the lensed arc is also detected.
This suggests that we detect emission in both
the foreground and the background galaxy and that the
emission is marginally resolved along its major axis.
We subtract a point source model in the visibility-plane to remove the unresolved part of the
emission, which we here assume to be dominated by the foreground galaxy. The emission
in the residual map coincides spatially with the lensed arc. We measure a flux density 
of $S_{\nu}$\eq0.39\,$\pm$\,0.08\,mJy for this residual component.
This flux density is consistent with
the difference between the integrated and the peak flux density measured in the
original continuum map ($\sim$0.4 mJy). 
We therefore adopt $S_{\nu}$\eq0.39\pmm0.12 mJy as the best estimate for the 2\,mm continuum flux of 
the background galaxy (RXJ1131). 
We here quote a conservative uncertainty, which is derived by adding those associated with the flux density of the 
point-source model ($\delta S_{\nu}$\eq0.088\,mJy) and 
the peak flux in the residual map (0.08\,mJy) 
in quadrature. We caution that this does not account for that in the 
de-blending procedure, where we have assigned 100\% of the point source flux to the foreground galaxy.
We report the peak flux in the original map
($S_{\nu}$\,=\,799\pmm88\,$\mu$Jy\,beam\pmOne) for the foreground galaxy, which
is the best estimate possible at the resolution of our observations, but we acknowledge that a non-negligible contribution from the background source to the peak flux cannot be ruled out.

@@

The VLA C-band continuum image in \Fig{cont} shows resolved emission from the
jets and the core of the foreground elliptical galaxy
as well as emission toward the background quasar.
Multiple peaks are seen along the arc with their centroids
coincident with the optical emission from the quasar.
We extract the flux densities for the lensing arc and the radio core in \Tab{photometry}.
We find a spectral index of $\alpha^{\rm 2mm}_{\rm 6cm}$\,=\,$-$0.02\pmm0.07
for the foreground galaxy and $\alpha^{\rm 2mm}_{\rm 6cm}$\,=\,$-$0.35\pmm0.21 % differential
for the background galaxy by fitting
power-laws ($S_\nu \propto \nu^{\alpha}$) to their continuum fluxes at
5\,GHz and 2\,mm. 
The spectral slope derived for the background source is flatter than the typical slope of pure synchrotron emission \citep[$\alpha$\ssim$-$0.7; e.g.,][]{Andreani93a}. This likely suggests % {\bf DR: 3C273 flat-spectrum radio source BLAH}
that a fraction of the observed 2\,mm emission arises from thermal dust  emission. 
This spectral slope would be even shallower 
if the background source contributes to the unresolved fraction of the
2\,mm flux. This would also 
explain the flat spectrum found for the foreground galaxy, which is atypical of elliptical galaxies. 
Assuming a synchrotron slope of $\alpha$\ssim$-$0.7, we would expect
a flux density of $S_{\rm 2mm}$\eq0.122\pmm0.004 mJy at 2\,mm      % (139.256/4.8815)**-0.7 *1.273 = 0.122; error: 1.273 +/- 0.042 --> 0.042/1.273 * 0.122 = 0.004
for the background source due to synchrotron radiation,
and thus a flux excess of $S_{\rm 2mm}$\eq0.27\pmm0.08\,mJy relative to this expectation
likely arises due to thermal dust emission.

@@

% the companion
A better fit is found for the lens model of
the red-most channel if we add a second source component (see
top left panel in \Fig{model}). This is consistent with previous results
reported by \citet[hereafter B08]{Brewer08a}, who find an optically faint companion
(component F in their paper) $\sim$2.4\,kpc in projection from the AGN host galaxy in $V$-band,
and with \citetalias{Claeskens06a}, who find evidence for an interacting galaxy near RXJ1131.
Spatially, the red velocity component of the CO emission
also coincides with this component F. It is therefore likely that we
detect \bco emission in a companion galaxy.
\defcitealias{Brewer08a}{B08}

@@

Previous studies of RXJ1131 find evidence for differential lensing across 
the {\it HST} $V$-, $I$-, and $H$-bands, where the
magnification factor varies from 10.9 to 7.8 \citepalias{Claeskens06a}.
This indicates that the spatial positions and extents of different stellar populations 
within the host galaxy are dissimilar with respect to the caustic which may be a result of dust extinction and reddening.
The best-fit lens models obtained here for different CO channel slices show that differential lensing also plays 
an important role in the observed \bco emission, with a magnification factor $(\mu_{\rm L})$ that varies
from 8.7 to 3.1 across different kinematic components (\Tab{model}).
The asymmetry in the line profile is therefore predominantly a result of differential lensing, where 
the redshifted CO-emitting gas is only apparently much brighter than the
blueshifted component due to a higher magnification factor.
A secondary effect is likely a result of including emission from the companion in the most redshifted velocity channels.
The variation in $\mu_{\rm L}$ found across channels is consistent with the source plane
positions relative to the caustics in \Fig{model}, where the red wing
emission mainly originates near the cusp
of the caustic and the blue wing emission is located beyond the caustics.
In fact, the intrinsic line flux of the redshifted and
blueshifted emission in RXJ1131 (after subtracting a contribution from the companion)
are $I_{\small \bco}$\,=\,1.26\pmm 0.23 Jy\,\kms and 1.25\pmm0.23 Jy\,\kms, respectively,
implying an intrinsically symmetric line profile (\Fig{delensed}). This is consistent with the symmetric source-plane
velocity gradient in our lens model (\Fig{model} and \Fig{PV}).
% The difference in magnification factors between CO and optical-to-NIR
% may further indicate that the CO emitting region is more extended and/or
% has a spatial offset from the optical-to-NIR emission
% (\ie different positions and alignment relative to the caustics) given
% the lower magnification factor found for CO (median $\mu_{\rm L}$ = 5.5)
% than in NIR, while their spatial extents appear to be similar. But then it could just be because of the low resolution and
% the degeneracy between the intrinsic source size and the magnification factor.

@@

\begin{figure*}[!htbp]
\centering
\includegraphics[trim=0 0 30 0, clip, width=0.4\textwidth]{../Figures/veloGradient_markers}
\includegraphics[width=0.525\textwidth]{../Figures/bestfit_PV.eps}
\caption{Velocity gradient of the \bco emission observed in RXJ1131 and the de-lensed velocity gradient obtained from lens modeling.
Left: source-plane positions obtained from the best-fit lens models (presented in \Sec{lensmodel}) are shown 
as markers color-coded by their velocities
atop the observed first moment map (i.e., in the image-plane; see \Fig{CO21mom}). 
The contours are in steps of 50\,\kms. 
Despite the distorted first moment map in the image-plane due to differential lensing and beam smearing effects, the source-plane velocity gradient reconstructed from dynamical lens modeling suggests that
RXJ1131 is intrinsically a kinematically-ordered disk.
Right: PV slice extracted along the fitted major axis at PA\,=\,121\degr based on 
the reconstructed source-plane velocity gradient.
Dashed line shows the best-fit ``rotation curve'' using an arctangent model.
The vertical error bars show the channel width for
each model and the horizontal error bars are the
1$\sigma$ uncertainties on the source-plane positions along the major axis.
 \label{fig:PV}}
\end{figure*}

@@

The observed velocity offset between the BLR and NLR lines may be explained by
a recoiling black hole (BH), 
where the BLR is moving at high velocity relative to the bulk of its host galaxy
\citep{Madau04a, Bonning07a, Loeb07a}.
A recoiling black hole is expected
when a pair of uneven mass BH coalesce, during which their orbital energy is being released
as gravitational wave and a non-zero net angular momentum is being carried away.
Depending on their initial conditions (e.g., black hole mass ratio, spin-orbit orientation, spin magnitude), 
numerical relativity simulations have shown that recoil velocities can reach up to 
$v_{\rm kick}$\ssim4000\,\kms for spinning BHs, 
with typical recoil velocities of $v_{\rm kick}$\ssim100\,$-$\,500\,\kms 
\citep[\eg][]{Libeskind06a,Campanelli07a}.
Several sources have been proposed as recoiling BH candidates \citep[][]{Komossa08a, Civano10a, Steinhardt12a}.
% For the candidate SDSS J0927+2943 at $z$\,$\sim$\,0.713 \citep{Komossa08a}, 
However, \citet{Decarli14a} have recently refuted 
such scenario for one of the candidates --- SDSS J0927+2943 --- by finding that the redshift of its
BLR lines is indeed consistent with its CO systemic redshift.
This is in contrast with RXJ1131, where our CO observations confirm 
the redshifted BLR lines compared to the CO systemic redshift.
Since this scenario requires a coalesced BH, it would imply that RXJ1131 
is a product of a previous merger, which is not implausible and 
might also explain the highly spinning BH in RX1131 \citep[$a$\ssim0.9;][]{Reis14a}.

@@

The source-plane distribution of the gas and stellar populations of different ages
indicates that the CO gas is of similar spatial extent as the old and long-lasting stellar populations,
whereas regions of recent \SF may be embedded within the molecular gas reservoir as a result of
gas accumulation driven by interactions with the companion.
Based on dynamical mass constraints, we cannot rule out the possibility that the
compact \SF in the host galaxy
may be heavily dust-obscured.
Hence, the true extent of recent \SF may be as extended as the molecular gas
reservoir.
While properties such as CO linewidth, SFR, and gas mass found in RXJ1131
are consistent with those of local ULIRGs and high-$z$ starburst galaxies,
its SFE is comparable to those of nearby and high-$z$ disk galaxies rather than
\SB systems. This is in good agreement with its disk-like kinematic signatures and its extended molecular gas distribution.
We find a specific \SF rate (sSFR\ssim4 Gyr\pmOne) that is $\lesssim$1.5$\sigma$ higher than those of ``main sequence'' galaxies.
The slight elevation in sSFR over the main sequence suggests that
the on-going star formation activity in RXJ1131 could be enhanced by interactions with the companion.
Recent simulations have illustrated that the disk component of a gas-rich
progenitor galaxy with low SFE can be
retained upon merging since the efficiency at removing angular momentum of the gas via
gravitational torques provided by stellar components is reduced in such a system \citep{Springel05a, Robertson06a, Hopkins09a}.
As such, the extended gas disk of RXJ1131 together with its low SFE may indicate
that the \SF in RXJ1131 could form a
larger stellar bulge in the merger remnant upon coalescing.

>>>>

We detect 2.2\,mm continuum emission at an
 integrated flux density of
1.2\pmm0.2 mJy, with a peak flux of
$S_{\nu}$\,=\,799\pmm88\,$\mu$Jy\,beam\pmOne
centered on the lensing galaxy (\Fig{cont}).
% We note that the reported flux uncertainties do not include those from absolute flux calibration.
Slightly extended emission
along the lensed arc is also detected.
This suggests that we detect emission in both
the foreground and the background galaxy and that the
emission is marginally resolved along its major axis.
We subtract a point source model in the visibility-plane to remove the unresolved part of the
emission, which we here assume to be dominated by the foreground galaxy. The emission
in the residual map coincides spatially with the lensed arc. We measure a flux density 
of $S_{\nu}$\eq0.39\,$\pm$\,0.08\,mJy for this residual component.
This flux density is consistent with
the difference between the integrated and the peak flux density measured in the
original continuum map ($\sim$0.4 mJy). 
We therefore adopt $S_{\nu}$\eq0.39\pmm0.12 mJy as the best estimate for the 2\,mm continuum flux of 
the background galaxy (RXJ1131). 
We here quote a conservative error bar, which is derived by adding the uncertainty 
associated with the flux density of the 
point-source model ($\delta S_{\nu}$\eq0.088\,mJy) with
that of the peak flux in the residual map (0.08\,mJy) 
in quadrature. We caution that this does not account for the systematic uncertainties of the 
de-blending procedure, where we have assigned 100\% of the point source flux to the foreground galaxy.
We report the peak flux in the original map
($S_{\nu}$\,=\,799\pmm88\,$\mu$Jy\,beam\pmOne) for the foreground galaxy, which
is the best estimate possible at the resolution of our observations, but we acknowledge that a non-negligible contribution from the background source to the peak flux cannot be ruled out.

@@

The VLA C-band continuum image in \Fig{cont} shows resolved emission from the
jets and the core of the foreground elliptical galaxy
as well as emission toward the background quasar.
Multiple peaks are seen along the arc with their centroids
coincident with the optical emission from the quasar.
We extract the flux densities for the lensing arc and the radio core in \Tab{photometry}.
We find a spectral index of $\alpha^{\rm 2mm}_{\rm 6cm}$\,=\,$-$0.02\pmm0.07
for the foreground galaxy and $\alpha^{\rm 2mm}_{\rm 6cm}$\,=\,$-$0.35\pmm0.21 % differential
for the background galaxy by fitting
power-laws ($S_\nu \propto \nu^{\alpha}$) to their continuum fluxes at
5\,GHz and 2\,mm. 
The spectral slope derived for the background source is flatter than the typical slope of pure synchrotron emission \citep[$\alpha$\ssim$-$0.7; e.g.,][]{Andreani93a}. This likely suggests % {\bf DR: 3C273 flat-spectrum radio source BLAH}
that a fraction of the observed 2\,mm emission arises from thermal dust  emission. 
This spectral slope would be even shallower 
if the background source contributes to the unresolved fraction of the
2\,mm flux. This would also 
explain the flat spectrum found for the foreground galaxy, which is atypical of elliptical galaxies. 
Assuming a synchrotron slope of $\alpha$\ssim$-$0.7, we would expect
a flux density of $S_{\rm 2mm}$\eq0.122\pmm0.004 mJy at 2\,mm      % (139.256/4.8815)**-0.7 *1.273 = 0.122; error: 1.273 +/- 0.042 --> 0.042/1.273 * 0.122 = 0.004
for the background source due to synchrotron radiation,
and thus an excess of $S_{\rm 2mm}$\eq0.27\pmm0.08\,mJy relative to this expectation
likely arises due to thermal dust emission.

@@

% the companion
A better fit is found for the lens model of
the red-most channel if we add a second source component (see
top left panel in \Fig{model}). This is consistent with previous results
reported by \citet[hereafter B08]{Brewer08a}, who find an optically faint companion
(component F in their paper) $\sim$2.4\,kpc in projection from the AGN host galaxy in $V$-band,
and with \citetalias{Claeskens06a}, who find evidence for an interacting galaxy near RXJ1131.
Spatially, the red velocity component of the CO emission
also consistent with this component F. It is therefore likely that we
detect \bco emission in a companion galaxy.
\defcitealias{Brewer08a}{B08}

@@

Previous studies of RXJ1131 find evidence for differential lensing across 
the {\it HST} $V$-, $I$-, and $H$-bands, where the
magnification factor varies from 10.9 to 7.8 \citepalias{Claeskens06a}.
This indicates that emission from different stellar populations 
within the host galaxy are dissimilar with respect to the caustic. 
%This imply that different stellar populations in the host galaxy are not co-spatial, and have  
%different spatial extents, or
%may be a result of dust extinction and reddening.
The best-fit lens models obtained here for different CO channel slices show that differential lensing also plays 
an important role in the observed \bco emission, with a magnification factor $(\mu_{\rm L})$ that varies
from 8.7 to 3.1 across different kinematic components (\Tab{model}).
The asymmetry in the line profile (\Fig{co21spec} and \Fig{delensed}) is therefore predominantly a result of
the redshifted CO-emitting gas being more strongly-magnified 
than the 
blueshifted component.
A secondary effect is likely a result of including emission from the companion in the most redshifted velocity channels.
The variation in $\mu_{\rm L}$ found across channels is consistent with the source plane
positions relative to the caustics in \Fig{model}, where the red wing
emission mainly originates near the cusp
of the caustic and the blue wing emission is located beyond the caustics.
In fact, the intrinsic line flux of the redshifted and
blueshifted emission in RXJ1131 (after subtracting a contribution from the companion)
are $I_{\small \bco}$\,=\,1.26\pmm 0.23 Jy\,\kms and 1.25\pmm0.23 Jy\,\kms, respectively,
implying an intrinsically symmetric line profile (\Fig{delensed}). This is consistent with the symmetric source-plane
velocity gradient in our lens model (\Fig{model} and \Fig{PV}).
% The difference in magnification factors between CO and optical-to-NIR
% may further indicate that the CO emitting region is more extended and/or
% has a spatial offset from the optical-to-NIR emission
% (\ie different positions and alignment relative to the caustics) given
% the lower magnification factor found for CO (median $\mu_{\rm L}$ = 5.5)
% than in NIR, while their spatial extents appear to be similar. But then it could just be because of the low resolution and
% the degeneracy between the intrinsic source size and the magnification factor.

@@

\begin{figure*}[!htbp]
\centering
\includegraphics[trim=0 0 30 0, clip, width=0.4\textwidth]{../Figures/veloGradient_markers}
\includegraphics[width=0.525\textwidth]{../Figures/bestfit_PV.eps}
\caption{Velocity gradient of the \bco emission observed in RXJ1131 and the de-lensed velocity gradient obtained from lens modeling.
Left: source-plane positions obtained from the best-fit lens models (presented in \Sec{lensmodel}) are shown 
as markers 
atop the observed first moment map (i.e., in the image-plane; see \Fig{CO21mom}). 
The markers are color-coded by their centroid velocities.
The contours are in steps of 50\,\kms. 
Despite the distorted first moment map in the image-plane due to differential lensing and beam smearing effects, the source-plane velocity gradient reconstructed from dynamical lens modeling suggests that
RXJ1131 is intrinsically a kinematically-ordered disk.
Right: PV slice extracted along the fitted major axis at PA\,=\,121\degr based on 
the reconstructed source-plane velocity gradient.
Dashed line shows the best-fit ``rotation curve'' using an arctangent model.
The vertical error bars show the channel width for
each model and the horizontal error bars are the
1$\sigma$ uncertainties on the source-plane positions along the major axis.
 \label{fig:PV}}
\end{figure*}

@@

The observed velocity offset between the BLR and NLR lines may be explained by
a recoiling black hole (BH), 
where the BLR is moving at high velocity relative to the bulk of its host galaxy
\citep{Madau04a, Bonning07a, Loeb07a}.
Depending on the initial conditions of the black hole pair (e.g., black hole mass ratio, spin-orbit orientation, spin magnitude), 
numerical relativity simulations have shown that recoil velocities can reach up to 
$v_{\rm kick}$\ssim4000\,\kms for spinning BHs, 
with typical recoil velocities of $v_{\rm kick}$\ssim100\,$-$\,500\,\kms 
\citep[\eg][]{Libeskind06a,Campanelli07a}.
Several sources have been proposed as recoiling BH candidates \citep[][]{Komossa08a, Civano10a, Steinhardt12a}.
% For the candidate SDSS J0927+2943 at $z$\,$\sim$\,0.713 \citep{Komossa08a}, 
However, \citet{Decarli14a} have recently refuted 
such scenario for one of the candidates --- SDSS J0927+2943 --- by finding that the redshift of its
BLR lines is indeed consistent with its CO systemic redshift.
This is in contrast with RXJ1131, where our CO observations confirm 
the redshifted BLR lines compared to the CO systemic redshift.
Since this scenario requires a coalesced BH, it would imply that RXJ1131 
is a product of a previous merger, which is not implausible and 
might also explain the highly spinning BH in RX1131 \citep[$a$\ssim0.9;][]{Reis14a}.

@@

The source-plane distribution of the gas and stellar populations of different ages
indicates that the CO gas is of similar spatial extent as the old and long-lasting stellar populations,
whereas regions of recent \SF may be embedded within the molecular gas reservoir as a result of
gas accumulation driven by interactions with the companion.
Based on dynamical mass constraints, we cannot rule out the possibility that the
compact \SF in the host galaxy
may be heavily dust-obscured.
Hence, the true extent of recent \SF may be as extended as the molecular gas
reservoir.
While properties such as CO linewidth, SFR, and gas mass found in RXJ1131
are consistent with those of local ULIRGs and high-$z$ starburst galaxies,
its SFE is comparable to those of nearby and high-$z$ disk galaxies rather than
\SB systems. This is in good agreement with its disk-like kinematic signatures and its extended molecular gas distribution.
We find a specific \SF rate (sSFR\ssim4 Gyr\pmOne) that is $\lesssim$1.5$\sigma$ higher than those of ``main sequence'' galaxies.
The slight elevation in sSFR over the main sequence suggests that
the on-going star formation activity in RXJ1131 could be enhanced by interactions with the companion.
Recent simulations have illustrated that the disk component of a gas-rich
progenitor galaxy with low SFE can be
retained upon merging since the efficiency at removing angular momentum of the gas via
gravitational torques provided by stellar components is reduced in such a system \citep{Springel05a, Robertson06a, Hopkins09a}.
As such, the extended gas disk of RXJ1131 together with its low SFE may indicate
that the \SF in RXJ1131 could form a
larger stellar bulge in the remnant disk galaxy upon coalescing.
